---
title: "Statistical Inference Project Part 1"
author: "Azmi Mohamed Ridwan"
date: "March 27, 2016"
output: word_document
---

## Overview

In this project, the exponential distribution is investigated and compared to the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. lambda  will  be set to 0.2 for all of the simulations. The target of study is the distribution of averages of 40 exponentials which will be simulated a thousand times.

**The knitr to pdf is not working for R version installed on my machine. I have knitted to Word first before creating a PDF which resulted in some a spaces in the final document. The result is approximately 5 pages which is within the 3+3 pages allowed in the instructions** 

## Simulations
*Include English explanations of the simulations you ran, with the accompanying R code. Your explanations should make clear what the R code accomplishes.*

The first code snippet sets up the necessary basic variables for the simultation.
```{r}
set.seed(2016)
lambda <- 0.2
n <- 40
num.sim <- 1000

```

Using the exponential distribution function *rexp*, a matrix is created with 1000 rows with the columns containing the 40 exponentials.
```{r}
sim.matrix <- matrix(rexp(num.sim * n, rate = lambda), num.sim, n)

```

The mean for each row is calculated and a histogram is plotted.
```{r}
sim.mean <- rowMeans(sim.matrix)
hist(sim.mean, xlab="Mean of 40 exponentials", ylab= "Frequency", col="light blue", main="Histogram for Simulation")
```

## Sample Mean versus Theoretical Mean
*Include figures with titles. In the figures, highlight the means you are comparing. Include text that explains the figures and what is shown on them, and provides appropriate numbers.*

The mean of the simulation and the theoretical mean are calculated here. The simulation mean values is then added to the histogram
```{r}
actual.mean = mean(sim.mean)
theoretical.mean = 1/lambda
print(c(actual.mean, theoretical.mean))

hist(sim.mean, xlab="Mean of 40 exponentials", ylab= "Frequency", col="light blue", main="Histogram for Simulation")
abline(v=actual.mean, lwd="4", col="red")
text(7, 120, paste("Actual mean = ", round(actual.mean,3), "\n Theoretical mean = 5" ), col="black")

```
The mean of the simulation is very close to the theoretical mean.

## Sample Variance versus Theoretical Variance
*Include figures (output from R) with titles. Highlight the variances you are comparing. Include text that explains your understanding of the differences of the variances.*

The actual and theoretical variance calculations are:
```{r}
actual.var<- var(sim.mean)
theoretical.var <- (1/lambda)^2/n

```
The actual variance is **`r actual.var`** and the theoretical variance is **`r theoretical.var`**. Again both values are quite close.

## Distribution
*Via figures and text, explain how one can tell the distribution is approximately normal.*

We compare the distribution of the mean with a theoretical normal distribution. The histogram is replotted as probability densities and increased breakpoints. The smoothed density line for the simulation mean (**blue**) is overlaid together with the theoretical normal distribution (**green**).
```{r}
hist(sim.mean, freq = FALSE, breaks = 20, xlab="Mean of 40 exponentials", ylab= "Probablity Densities", col="light blue", main="Histogram for Simulation")
x <- c(0,1)
lines(density(sim.mean), lwd=3, col="blue")
curve(dnorm(x, mean=theoretical.mean, sd=sqrt(theoretical.var)), col="green", lwd=2, add=TRUE, yaxt="n")

```
We can see that the simulation distribution is very close to the theoretical normal distribution.

We can do a Quantile-Quantile plot. We can see the values of the simulations fall close on the diagonal.

```{r}
qqnorm(sim.mean, main="Q-Q Plot for Simulation", xlab="Theoretical Quantiles", ylab="Sample Quantiles")
qqline(sim.mean, col="blue")
```

We can also compare the confidence interval. Again, we can see that the 2 values are very close. 
```{r}
actual.ci <- round (mean(sim.mean) + c(-1,1)*1.96*sd(sim.mean)/sqrt(n),3)
theoretical.ci <- theoretical.mean + c(-1,1)*1.96*sqrt(theoretical.var)/sqrt(n)
```
95% confidence interval - **Actual=[`r actual.ci`]**, **Theoretical=[`r theoretical.ci`]**


